---
title: "Exploratory Data Analysis"
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(RSocrata) # for reading in data
library(lubridate) # for dealing with date-time data
```

# Background

We're going to explore a large data set or traffic crashes to learn about what factors are connected with injuries. We will use data from the city of [Chicago's open data portal](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if/about_data). (This activity is derived from a [blog post](https://juliasilge.com/blog/chicago-traffic-model/) by Julia Silge)

```{r}
years_ago <- mdy("01/01/2022") # data from last 2 years. May take time to load!
crash_url <- glue::glue("https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if?$where=CRASH_DATE >= '{years_ago}'")
crash_raw <- as_tibble(read.socrata(crash_url)) # a new way to read in data, don't worry about it!
```

This dataset is pretty crazy! Take a look at it in the viewer, and then let's do some data munging to get it into a nicer form. 

  -create a variable called `injuries` which indicates if the crash involved injuries or not.
  -create an unknown category for missing `report_type`s
  -decide which other variables to keep
  
```{r}
crash <- crash_raw %>%
  arrange(desc(crash_date)) %>%
  transmute(
    injuries = as.factor(if_else(injuries_total > 0, "injuries", "none")),
    # choose your variables here (use ?transmute to see what this does)
    latitude, longitude
  )
```


## Exploratory Data Analysis

Here's a few questions to get you started. 

1. Take a look at crashes by latitude and longitude, colored by injuries. What do you notice?

2. What are the most common contributing factors to a crash?

3. How do crashes vary month by month? Compare crashes by month in 2022 to 2023.

4. Are crashes more likely to cause injuries when it is rainy and dark? Use the variables `weather_condition` and `lighting_condition` to explore.

5. Choose a question you want to explore, and create an appropriate visual.


