---
title: "R Lab #"
author: "Your Name"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#This is a setup chunk, you will never need to modify it
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openintro)
library(oilabs)
```

### Exercise 1

Here you might type any answers to Exercises.

```{r ex1}
#Here you would type any code for Exercise 1
```

### Exercise 2

Here you might type any answers to Exercises.

```{r ex2}
#Here you would type any code for Exercise 2
```


### Exercise 3

Here you might type any answers to Exercises.

```{r ex3}
#Here you would type any code for Exercise 3
```

### Exercise 4

Here you might type any answers to Exercises.

```{r ex4}
#Here you would type any code for Exercise 4
```

### Exercise 5

Here you might type any answers to Exercises.

```{r ex5}
#Here you would type any code for Exercise 5
```

### Exercise 6

Here you might type any answers to Exercises.

```{r ex6}
#Here you would type any code for Exercise 6
```

### Exercise 7

Here you might type any answers to Exercises.

```{r ex7}
#Here you would type any code for Exercise 7
```

### Exercise 8

Here you might type any answers to Exercises.

```{r ex8}
#Here you would type any code for Exercise 8
```

### Exercise 9

Here you might type any answers to Exercises.

```{r}
#Here you would type any code for Exercise 9
```


